Generalized Chacon polynomial constructions

### 2018-02-19 article.tex
intro and bibliography in article

\usepackage{bbm}

\begin{document}
\section{Basic definitions}
\subsection{The polynomials $P_m^p$ in the p-adic group}
Let $p \ge 3$ be an arbitary integer.
Consider the compact group of $p$-adic numbers
$$\Gamma := \mathbb{Z}_p = \left\{x = \left(x_0, x_1, x_2, \ldots \right), x_k \in \{0, 1, \ldots, p - 1\} \right\}$$
We will also use the set
$$\Gamma' := \Gamma \setminus \{(p-1,p-1,\ldots)\},$$
where each element has only a finite number of leading elements equal to $(p-1)$.

@@

\section{Palindromic property}
In this section, we are going to prove that, just like in the classical case, the polynomials $P_m^p$ are palindromic for any parameter $p$. Moreover, we generalize the definition of $\phi$ functional and describe the class of Chacon-like functionals which produce palindromic polynomials.

@@

To prove the theorem, it is now enough to notice that in Definition 1 for $\phi$ produced by binary-antipalindromic $\omega$ we may substitute $\phi_\star = 0, \phi^\star = 1, \zeta=1$ and hence the palindromic properly is equivalent to $\forall j, 0 \le j \le m:\ \pi_m(j)=\pi_m(m-j)$.
This summarizes the consideration of $\omega$ functions with $\mathrm{Ran }\ \omega = \{0,1\} $. Let us skip to the general construction.

@@

\section{Recurrence formulae for linear $\omega$}
\begin{notation} We denote $n$th triangle number $\frac{n(n+1)}{2}$ as $\Delta_n$ \end{notation}
\begin{lemma}\label{simpleCase} $P_{pm}^p(t)=t^{m\Delta_{p-2}}P_m^p(t)$
\end{lemma}
\begin{proof}
Let $x \in \Gamma'$. Recalling the structure of $\{\phi(S^j x)\}_{j \ge 0}$, the value of $\phi^{(pm)}(x)$ is the sum of:
\begin{itemize}
\item the order-$0$ points. There are exactly $(p-1)m$ points following the repeating pattern $\ldots,(p-2),0,1,2,...,(p-2),0,1,\ldots$. Their contribution is $m$ times the sum of integers $0,1,\ldots,(p-2)$ which is $\frac{(p-1)(p-2)}{2}m=m\Delta_{p-2}$
\item the higher-order points. Their contribution is $\phi^{(m)}(\sigma x)$.
\end{itemize}
Hence $\phi^{pm}(x) = m\Delta_{p-2} + \phi^{(m)}(\sigma x)$. By the definition of polynomials $P_m(t)$ it implies 
$P_pm(t) = \mathbb{E}_\lambda\left[ t^{\phi^{(pm)}(x)}\right] = 
\mathbb{E}_\lambda\left[ t^{m\Delta_{p-2} + \phi^{(m)}(\sigma x)}\right] = 
t^{m\Delta_{p-2}} \mathbb{E}_\lambda \left[ t^{\phi^{(m)} (\sigma x)}  \right] = 
t^{m\Delta_{p-2} }P_m(t)$
\end{proof}
\begin{notation}
Similarly to $\phi^{(m)}$, we denote $\omega^{(m)}(j) := \omega(j) + \omega(j+1) + \ldots + \omega(j+m-1)$ if $j + m < p - 1$.
\end{notation}
\begin{proposition} $\omega^{(k)}(j) = k j + \Delta_{k-1}$
\end{proposition}
\begin{proof}
The value of $\omega^{(k)}(x_0)$ is the sum of an arithmetic progression $x_0, x_0 + 1, \ldots, x_0 + k-1$.
\end{proof}
\begin{lemma} \label{phiLemma}
Let $0 < k < p$. Consider $x=(x_0,x_1,\ldots) \in \Gamma'$.

@@

We substitute the index $j$ with $q = p-1-j$ and simplify the exponent:
$$  (p-1-j) j + \Delta_{p-2-j} + \Delta_{j+k-p-1} =
    q(p-1-q) + \Delta_{q-1} + \Delta_{k-2-q}=
   q(p-k)+\Delta_{k-2} 
$$   
Now we collect the result: 
\begin{equation}\label{eq:secondCondExp}
\mathbb{E}_\lambda\left[ t^{\phi^{(pm+k)}(x)}| x_0 \ge p-k \right] 
 =  \frac{1}{k} t^{m\Delta_{p-2} + \Delta_{k-2}} P_{m+1}^p(t)
  \sum\limits_{q=0}^{k-1} t^{q(p-k)}
  \end{equation}
It remains to substitute (\ref{eq:firstCondExp}) and (\ref{eq:secondCondExp}) into (\ref{eq:totalExp}) to obtain the recurrent equation:
$$P_{pm+k}(t)  =   \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-1}}\sum\limits_{j=0}^{p-k-1} t^{jk} P_m^p(t) 
      + \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-2}}\sum\limits_{j=0}^{k-1} t^{j(p-k)} P_{m+1}^p(t) $$
\end{proof}
Bringing together the results of \Cref{simpleCase} and \Cref{complexCase}, we obtain the general recurrent law for $P_{n}^p(t)$
\begin{theorem}[Recurrent formulae on $P_{n}^p(t)$] 
\begin{align}
\label{P_pm} P_{pm}^p(t) & =  t^{m\Delta_{p-2}}\ P_m^p(t) \\
\label{P_pm+k} P_{pm+k}^p(t) & =  \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-1}}\sum\limits_{j=0}^{p-k-1} t^{jk} P_m^p(t)  + \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-2}}\sum\limits_{j=0}^{k-1} t^{j(p-k)} P_{m+1}^p(t), \ 0 < k < p
 \end{align}
\end{theorem}

\section{Degrees of polynomials}
\subsection{Sequence $\hat{s}_j$}
Here we define the substitution-produced sequence $\{\hat{s}_j\}_{j \ge 0}$. It will be useful to evaluate the degrees of $P_m^p(t)$. We build this sequence recursively. First, fix the parameter value $p \ge 3$.
\begin{definition}[Zero-tier construction]
Let $\hat{s}_0^{(0)} = p-2,\ \hat{s}_j^{(0)} = p-2-(j-1)$ for $j = \overline{1, p-1}$.
\end{definition}
Literally, $\{ \hat{s}_j^{(0)}\}$ is $p-2, p-2, p-3, \ldots, 2, 1, 0$.
\begin{definition} [Number substitution rule]
We substitute
\begin{itemize}
 \item $0$ with $p-2, p-3, \ldots, 2, 1, 0, 0$;
 \item $1$ with $p-2, p-3, \ldots, 2, 1, 1, 0$;
 \item $2$ with $p-2, p-3, \ldots, 2, 2, 1, 0$;
 \item \ldots
 \item $p-3$ with $p-2, p-3, p-3, \ldots, 2, 1, 0$;
 \item $p-2$ with $p-2, p-2, p-3, \ldots, 2, 1, 0$;

@@

\subsection{Recurrent formulae for degrees}
\begin{definition}
Let $m \ge 0$. We define $d_m^p := \mathrm{deg}\ P_m^p$.
\end{definition}
\begin{proposition} \label{dm_eq}
For any $m \ge 0,\ 0 < k < p$:
\begin{align*}
  d_{pm}^p &= m \Delta_{p-2} + d_m^p \\
  d_{pm+k}^p &= m \Delta_{p-2}+(p-k)(k-1)+\Delta_{k-2}+d_m^p + \mathrm{max}\{p-k-1, d_{m+1}^p-d_m^p\}
\end{align*}
\end{proposition}
\begin{proof}
Directly from the (\ref{P_pm}) and (\ref{P_pm+k}) we obtain the equations for degrees:
\begin{align*}
  d_{pm}^p &= m \Delta_{p-2} + d_m^p \\
  d_{pm+k}^p &= \mathrm{max}\{ m\Delta_{p-2} + \Delta_{k-1} + k(p-k-1) + d_m^p,\     
                                m\Delta_{p-2} + \Delta_{k-2} + (p-k)(k-1) + d_{m+1}^p \}
\end{align*}
To show this proposition, it is enough to take the common terms out of the maximum operator.
\end{proof}
The following holds for aribtary fixed $p \ge 3$. Keeping this in mind, we use $d_j$ instead of $d_j^p$ for short.
\begin{definition} \label{sm_def}
For any $m \ge 0$, we define $s_m := d_{m+1}^p - d_m^p $.
\end{definition}
Our goal is to show that this sequence $\{s_m\}_{m \ge 0}$ is equal to $\{\hat{s}_m\}$ defined in the previous Subsection. First, we deduce recurrent equations on $s_m$.
\begin{proposition}
\begin{equation} \label{sm_eq}
s_{pm+k} = p-k-1 - \mathbbm{1}\{s_m < p-1-k\}
\end{equation}
\end{proposition}
\begin{proof}
We will show this using \Cref{dm_eq} and \Cref{sm_def}. If we try to evaluate $s_j$, it collapses into two cases:
\begin{itemize}
\item Let $k<p-1 $. $$s_{pm+k} = d_{pm+k+1}-d_{pm+k} =$$ $$= m\Delta_{p-2} + (p-k-1)k + \Delta_{k-1} + d_m + \max\{p-k-1, d_{m+1}-d_m\} - $$ $$ - m\Delta_{p-2}-(p-k)(k-1)-\Delta_{k-2}-d_m-\max\{p-k,d_{m+1}-d_m\}=$$ $$ =p-k-1+\max\{p-k-1, d_{m+1}-d_m\}-\max\{p-k,d_{m+1}-d_m\};$$
We substitute $d_{m+1}-d_m = s_m$ and observe that $$\max\{p-k-1, d_{m+1}-d_m\}-\max\{p-k,d_{m+1}-d_m\}= \mathbbm{1}\{s_m < 1-k\}.$$ This subsitution leads to exaclty the equation in the proposition.
\item $s_{pm+p-1} = d_{p(m+1)}-d_{pm+3} = \\=(m+1)\Delta_{p-2}+d_{m+1}-m\Delta_{p-2}-(p-(p-1))(p-1-1)-\Delta_{p-1-2}-d_m-\max\{0, d_{m+1}-d_m\}=\\=\Delta_{p-2}+(d_{m+1}-d_m)-(p-2)-\Delta_{p-3}-(d_{m+1}-d_m)=0$
\end{itemize}

@@

\begin{lemma}\label{sm1}
For $j = 0, 1, \ldots, p^2-1:\ s_j=\hat{s}_j^{(1)}$
\end{lemma}
\begin{proof}
We compute the initial $p^2$ terms of this sequence. Recall that $P_0^p(t)=1,\ P_1^p(t) = \frac{1}{p-1} \sum\limits_{j=0}^{p-2} t^j$, hence $d_0 = 0,\ d_1 = p-2$ and $s_0 = p-2$. We will use this as the initial condition in the recurrent equation from (\ref{sm_eq}).
Observe the equation (\ref{sm_eq}). The right part consists of a periodic term $p-k-1$ with period $p$ and the step function $-\mathbbm{1}\{s_m < p-1-k\}$ which is determined by the number of period. Thus we evaluate the sequence $\{s_j\}$ period-wise.

@@

To prove the lemma, we will this equation. Let $m \ge p,\ 0 < k < p-1$. From (\ref{sm_eq}) we obtain $s_{pm+k}=p-k-1 - \mathbbm{1}\{s_m < p-1-k\}$ and $s_{p(p-1-s_m)+k}= p-k-1 - \mathbbm{1}\{s_{p-1-s_m} < p-1-k\}$. One may easily observe that the equality $s_m=p-1-s_m$ turns this into an identity.
\end{proof}
\begin{theorem}
For each $j \ge 0$: $s_j=\hat{s}_j$.
\end{theorem}
\begin{proof}
The results of \Cref{sm1} and \Cref{sm2} show that \Cref{sm_good_def} is true for $\{s_j\}_{j \ge 0}$.
\end{proof}
\begin{corollary}
$d_m = \sum\limits_{j=0}^{m-1} \hat{s}_j$
\end{corollary}
\end{document}

>>>>

\usepackage{bbm}
\bibliographystyle{unsrt}

\title{Generalized Chacon polynomial constructions}

\author{Vladislav Slyusarev}

\date{\today}

\begin{document}
\maketitle

\section{Introduction}
The classical version of Chacon’s automorphism, which is the main subject of the
present work, was described by Friedman in \cite{friedman}. It is a famous example of a rank-one automorphism, for which we recall briefly the construction by cutting and stacking:
We start with a Rokhlin tower of height $h_0 := 1$, called Tower 0. At step $n$, Tower
$n-1$ (of height $h_{n-1}$) is cut into $3$ sub-columns, a spacer is inserted above the middle
column before stacking all parts to get Tower $n$, of height $h_n=3h_{n-1}+1$. This
transformation, denoted hereafter by $T$ and acting on a standard Borel probability
space $(X,\mathcal{A},\mu)$, is known to present an interesting combination of ergodic and
spectral properties. It is weakly mixing but not strongly mixing (see \cite{mixing}, where
the historical version of Chacon’s automorphism is constructed with only $2$ subcolumns,
but whose arguments also apply in the classical case). Del Junco proved
in \cite{deljunco} that T has trivial centralizer, then improved this result by showing with
Rahe and Swanson that it has minimal self-joinings \cite{joinings}. Prikhod'ko and Ryzhikov proved in \cite{convolutions} that the convolution powers of its maximal spectral type are pairwise mutually singular. Their method involves the identification, in the weak
closure of the powers of the associated Koopman operator $\hat{t}$, of an infinite family
of polynomials in $\hat{T}$.

An automorphism $S$ is said to be \emph{$\alpha$-weakly mixing} (for some $0\le \alpha \le 1$) if there exists a sequence $(k_j)$ of integers such that $\hat{S}^{k_j}$ converges weakly to $\alpha \Theta + (1-\alpha)\mathrm{Id}$, where $\Theta$ is the ortho-projector to constants. The disjointness of the convolution powers is automatically satisfied in the case of $\alpha$-weakly mixing transformations with $0 < \alpha < 1$ (see Katok \cite{katok} and Stepin \cite{stepin}). This property has been applied for
numerous counterexamples in ergodic theory \cite{deljunco1992}. The question of $\alpha$-weak mixing for Chacon’s automorphism is a special case of a general problem to tell which operators
can be obtained as weak limits of powers of $\hat{T}$. For a recent application of weak
limits of powers of the Koopman operator, see \cite{delarue2012}. Examples of transformations
with non-trivial explicit weak closure of powers are given in \cite{ryzhikov}.

E. Janvresse, A. A. Prikhod’ko, T. de la Rue, and V. V. Ryzhikov have shown in \cite{weaklimits} that the weak closure of the powers of $\hat{T}$ is reduced to $\Theta$
and an explicit family of polynomials $P_m(\hat{T})$:

$$
\mathcal{L} = {\Theta} \cup \{P_{m_1}(\hat{T}),\ldots,P_{m_r}(\hat{T})\hat{T}^n,r\ge0,\ 1\le m_1\le\ldots\le m_r,\ n\in \mathbb{Z}\}.
$$

 These polynomials are obtained from the representation of $T$ as an integral automorphism over the $3$-adic odometer (this representation was also used earlier in \cite{convolutions}). Several properties and the recurrence formulae for these polynomials are also described in \cite{weaklimits}.

The purpose of the present paper is to generalize the results of \cite{weaklimits} connected to the family of polynomials. We consider a parametric set of transformations similar to the Chacon’s automorphism and infer the properties and the recurrence equations for the polynomials generated by these transformations. The constructions are parametrized by an integer $p \ge 3$, and the classic case of Chacon's automorphism corresponds to the parameter value $p=3$. In \Cref{definitions}, we give a definition of the functional $\phi$, the polynomials $P_m^p(t)$ and the generalized rautomorphism $T$. We provide
inductive formulas for these polynomials in \Cref{recurrence}. The proofs of several important properties of $P_m^p(t)$, such as the palindromic property and the sequence of degrees, are provided in \Cref{palindromic} and \Cref{degrees} respectively.

\section{Basic definitions}\label{definitions}
\subsection{The polynomials $P_m^p$ in the p-adic group}
Let $p \ge 3$ be an arbitary integer.
Consider the compact group of $p$-adic numbers
$$\Gamma := \mathbb{Z}_p = \left\{x = \left(x_0, x_1, x_2, \ldots \right), x_k \in \{0, 1, \ldots, p - 1\} \right\}$$
We will also use the set
$$\Gamma' := \Gamma \setminus \{(p-1,p-1,\ldots)\},$$
where each element has only a finite number of leading elements equal to $(p-1)$.

@@

\section{Palindromic property}\label{palindromic}
In this section, we are going to prove that, just like in the classical case, the polynomials $P_m^p$ are palindromic for any parameter $p$. Moreover, we generalize the definition of $\phi$ functional and describe the class of Chacon-like functionals which produce palindromic polynomials.

@@

@@

\section{Recurrence formulae for $P_m^p(t)$}\label{recurrence}
\begin{notation} We denote $n$th triangle number $\frac{n(n+1)}{2}$ as $\Delta_n$ \end{notation}
\begin{lemma}\label{simpleCase} $P_{pm}^p(t)=t^{m\Delta_{p-2}}P_m^p(t)$
\end{lemma}
\begin{proof}
Let $x \in \Gamma'$. Recalling the structure of $\{\phi(S^j x)\}_{j \ge 0}$, the value of $\phi^{(pm)}(x)$ is the sum of:
\begin{itemize}
\item the order-$0$ points. There are exactly $(p-1)m$ points following the repeating pattern $\ldots,(p-2),0,1,2,...,(p-2),0,1,\ldots$. Their contribution is $m$ times the sum of integers $0,1,\ldots,(p-2)$ which is $\frac{(p-1)(p-2)}{2}m=m\Delta_{p-2}$
\item the higher-order points. Their contribution is $\phi^{(m)}(\sigma x)$.
\end{itemize}
Hence $\phi^{pm}(x) = m\Delta_{p-2} + \phi^{(m)}(\sigma x)$. By the definition of polynomials $P_m(t)$ it implies 
$P_pm(t) = \mathbb{E}_\lambda\left[ t^{\phi^{(pm)}(x)}\right] = 
\mathbb{E}_\lambda\left[ t^{m\Delta_{p-2} + \phi^{(m)}(\sigma x)}\right] = 
t^{m\Delta_{p-2}} \mathbb{E}_\lambda \left[ t^{\phi^{(m)} (\sigma x)}  \right] = 
t^{m\Delta_{p-2} }P_m(t)$
\end{proof}
\begin{notation}
Similarly to $\phi^{(m)}$, we denote $\omega^{(m)}(j) := \omega(j) + \omega(j+1) + \ldots + \omega(j+m-1)$ if $j + m < p - 1$.
\end{notation}
\begin{proposition} $\omega^{(k)}(j) = k j + \Delta_{k-1}$
\end{proposition}
\begin{proof}
The value of $\omega^{(k)}(x_0)$ is the sum of an arithmetic progression $x_0, x_0 + 1, \ldots, x_0 + k-1$.
\end{proof}
\begin{lemma} \label{phiLemma}
Let $0 < k < p$. Consider $x=(x_0,x_1,\ldots) \in \Gamma'$.

@@

We substitute the index $j$ with $q = p-1-j$ and simplify the exponent:
$$  (p-1-j) j + \Delta_{p-2-j} + \Delta_{j+k-p-1} =
    q(p-1-q) + \Delta_{q-1} + \Delta_{k-2-q}=
   q(p-k)+\Delta_{k-2} 
$$   
Now we collect the result: 
\begin{equation}\label{eq:secondCondExp}
\mathbb{E}_\lambda\left[ t^{\phi^{(pm+k)}(x)}| x_0 \ge p-k \right] 
 =  \frac{1}{k} t^{m\Delta_{p-2} + \Delta_{k-2}} P_{m+1}^p(t)
  \sum\limits_{q=0}^{k-1} t^{q(p-k)}
  \end{equation}
It remains to substitute (\ref{eq:firstCondExp}) and (\ref{eq:secondCondExp}) into (\ref{eq:totalExp}) to obtain the recurrence equation:
$$P_{pm+k}(t)  =   \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-1}}\sum\limits_{j=0}^{p-k-1} t^{jk} P_m^p(t) 
      + \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-2}}\sum\limits_{j=0}^{k-1} t^{j(p-k)} P_{m+1}^p(t) $$
\end{proof}
Bringing together the results of \Cref{simpleCase} and \Cref{complexCase}, we obtain the general recurrence law for $P_{n}^p(t)$
\begin{theorem}[Recurrence formulae on $P_{n}^p(t)$] 
\begin{align}
\label{P_pm} P_{pm}^p(t) & =  t^{m\Delta_{p-2}}\ P_m^p(t) \\
\label{P_pm+k} P_{pm+k}^p(t) & =  \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-1}}\sum\limits_{j=0}^{p-k-1} t^{jk} P_m^p(t)  + \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-2}}\sum\limits_{j=0}^{k-1} t^{j(p-k)} P_{m+1}^p(t), \ 0 < k < p
 \end{align}
\end{theorem}

\section{Degrees of polynomials}\label{degrees}
\subsection{Sequence $\hat{s}_j$}
Here we define the substitution-produced sequence $\{\hat{s}_j\}_{j \ge 0}$. It will be useful to evaluate the degrees of $P_m^p(t)$. We build this sequence recursively. First, fix the parameter value $p \ge 3$.
\begin{definition}[Zero-tier construction]
Let $\hat{s}_0^{(0)} = p-2,\ \hat{s}_j^{(0)} = p-2-(j-1)$ for $j = \overline{1, p-1}$.
\end{definition}
Literally, $\{ \hat{s}_j^{(0)}\}$ is $p-2, p-2, p-3, \ldots, 2, 1, 0$.
\begin{definition} [Number substitution rule]
We substitute
\begin{itemize}
 \item $0$ with $p-2, p-3, \ldots, 2, 1, 0, 0$;
 \item $1$ with $p-2, p-3, \ldots, 2, 1, 1, 0$;
 \item $2$ with $p-2, p-3, \ldots, 2, 2, 1, 0$;
 \item \ldots
 \item $p-3$ with $p-2, p-3, p-3, \ldots, 2, 1, 0$;
 \item $p-2$ with $p-2, p-2, p-3, \ldots, 2, 1, 0$;

@@

\subsection{Recurrence formulae for degrees}
\begin{definition}
Let $m \ge 0$. We define $d_m^p := \mathrm{deg}\ P_m^p$.
\end{definition}
\begin{proposition} \label{dm_eq}
For any $m \ge 0,\ 0 < k < p$:
\begin{align*}
  d_{pm}^p &= m \Delta_{p-2} + d_m^p \\
  d_{pm+k}^p &= m \Delta_{p-2}+(p-k)(k-1)+\Delta_{k-2}+d_m^p + \mathrm{max}\{p-k-1, d_{m+1}^p-d_m^p\}
\end{align*}
\end{proposition}
\begin{proof}
Directly from the (\ref{P_pm}) and (\ref{P_pm+k}) we obtain the equations for degrees:
\begin{align*}
  d_{pm}^p &= m \Delta_{p-2} + d_m^p \\
  d_{pm+k}^p &= \mathrm{max}\{ m\Delta_{p-2} + \Delta_{k-1} + k(p-k-1) + d_m^p,\     
                                m\Delta_{p-2} + \Delta_{k-2} + (p-k)(k-1) + d_{m+1}^p \}
\end{align*}
To show this proposition, it is enough to take the common terms out of the maximum operator.
\end{proof}
The following holds for aribtary fixed $p \ge 3$. Keeping this in mind, we use $d_j$ instead of $d_j^p$ for short.
\begin{definition} \label{sm_def}
For any $m \ge 0$, we define $s_m := d_{m+1}^p - d_m^p $.
\end{definition}
Our goal is to show that this sequence $\{s_m\}_{m \ge 0}$ is equal to $\{\hat{s}_m\}$ defined in the previous Subsection. First, we deduce recurrence equations on $s_m$.
\begin{proposition}
\begin{equation} \label{sm_eq}
s_{pm+k} = p-k-1 - \mathbbm{1}\{s_m < p-1-k\}
\end{equation}
\end{proposition}
\begin{proof}
We will show this using \Cref{dm_eq} and \Cref{sm_def}. If we try to evaluate $s_j$, it collapses into two cases:
\begin{itemize}
\item Let $k<p-1 $. $$s_{pm+k} = d_{pm+k+1}-d_{pm+k} =$$ $$= m\Delta_{p-2} + (p-k-1)k + \Delta_{k-1} + d_m + \max\{p-k-1, d_{m+1}-d_m\} - $$ $$ - m\Delta_{p-2}-(p-k)(k-1)-\Delta_{k-2}-d_m-\max\{p-k,d_{m+1}-d_m\}=$$ $$ =p-k-1+\max\{p-k-1, d_{m+1}-d_m\}-\max\{p-k,d_{m+1}-d_m\};$$
We substitute $d_{m+1}-d_m = s_m$ and observe that $$\max\{p-k-1, d_{m+1}-d_m\}-\max\{p-k,d_{m+1}-d_m\}= \mathbbm{1}\{s_m < 1-k\}.$$ This subsitution leads to exaclty the equation in the proposition.
\item $s_{pm+p-1} = d_{p(m+1)}-d_{pm+3} = \\=(m+1)\Delta_{p-2}+d_{m+1}-m\Delta_{p-2}-(p-(p-1))(p-1-1)-\Delta_{p-1-2}-d_m-\max\{0, d_{m+1}-d_m\}=\\=\Delta_{p-2}+(d_{m+1}-d_m)-(p-2)-\Delta_{p-3}-(d_{m+1}-d_m)=0$
\end{itemize}

@@

\begin{lemma}\label{sm1}
For $j = 0, 1, \ldots, p^2-1:\ s_j=\hat{s}_j^{(1)}$
\end{lemma}
\begin{proof}
We compute the initial $p^2$ terms of this sequence. Recall that $P_0^p(t)=1,\ P_1^p(t) = \frac{1}{p-1} \sum\limits_{j=0}^{p-2} t^j$, hence $d_0 = 0,\ d_1 = p-2$ and $s_0 = p-2$. We will use this as the initial condition in the recurrence equation from (\ref{sm_eq}).
Observe the equation (\ref{sm_eq}). The right part consists of a periodic term $p-k-1$ with period $p$ and the step function $-\mathbbm{1}\{s_m < p-1-k\}$ which is determined by the number of period. Thus we evaluate the sequence $\{s_j\}$ period-wise.

@@

To prove the lemma, we will this equation. Let $m \ge p,\ 0 < k < p-1$. From (\ref{sm_eq}) we obtain $s_{pm+k}=p-k-1 - \mathbbm{1}\{s_m < p-1-k\}$ and $s_{p(p-1-s_m)+k}= p-k-1 - \mathbbm{1}\{s_{p-1-s_m} < p-1-k\}$. One may easily observe that the equality $s_m=p-1-s_m$ turns this into an identity.
\end{proof}
\begin{theorem}
For each $j \ge 0$: $s_j=\hat{s}_j$.
\end{theorem}
\begin{proof}
The results of \Cref{sm1} and \Cref{sm2} show that \Cref{sm_good_def} is true for $\{s_j\}_{j \ge 0}$.
\end{proof}
\begin{corollary}
$d_m = \sum\limits_{j=0}^{m-1} \hat{s}_j$
\end{corollary}
\newpage
\bibliography{chacon}

\end{document}
